A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-11-18 kobakian-cook.Rmd
Earl's comments up to Algorithm section

---
title: A Hexagon Tile Map Algorithm for Displaying Spatial Data
author:
  # see ?rjournal_article for more information
  - name: Stephanie Kobakian
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  stephanie.kobakian@monash.edu 
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
  - name: Earl Duncan
    affiliation: Queensland University of Technology
    address:
    - School of Mathematical Sciences
    email:  earl.duncan@qut.edu.au
abstract: >
  The algorithm presented creates a display that uses tessellated hexagons to represent a set of spatial polygons. It allocates these hexagon in a manner that preserves the spatial relationship of the geographic units, in light of their positions to points of interest. The display showcases spatial distributions, by emphasising the small geographical regions that are often difficult to locate on geographic maps. Spatial distributions have been presented on alternative representations of geography for many years. In modern times, interactivity and animation have allowed alternative displays to play a larger role. Alternative representations have been popularised by online news sites, and atlas websites with a focus on public consumption. Applications are increasingly widespread, especially in the areas of disease mapping, and election results.
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: kobakian-cook.bib
output: rticles::rjournal_article
editor_options: 
  chunk_output_type: console
---

@@

Over time the density of living in major cities has increased as residents have gathered to live near urban areas [@ACTUC]. The division of these resident into approximately equal population areas results in dramatically different square meterage of administrative geographic units, such as states or electorates. In a choropleth map display the geographic units are filled with colour to represent the value of the statistic for each unit [@EI]. This can allow unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:choro} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for females, in each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s around Brisbane and in northern New South Wales stand out as more orange and red. However, this display neglects the vast amount of Australia residents living in the densely populated capital cities.

@@

The solutions to this visualisation problem begin with the geography. Alternative maps shift the focus from land area and shape, to the value of the statistics in a collection of areas [@ACCAC]. Each style of cartogram applies a different transformation to the geographic areas, to highlight the values of the statistic of interest. These displays result in a distortion of the map space to represent features of the distribution across the areas [@ACCAC] as the statistic of interest is used to determine the cartogram layout.

Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing the population in each administrative area [@TAAM]. This acknowledges that the number of residents can be different but recognises that each area, or person within it is equally important.

This paper contains a discussion of Existing Mapping Practices in Section 2. This is followed by details of the Algorithm in Section 3. Section 4 describes the implementation of the algorithm in the `sugarbag` package. Section 5 outlines the use and benefits of animation.

@@

<!-- choropleth -->
The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC]. The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In Figure \ref{fig:tas_displays} the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

There are several established alternative visualisation methods. Tile maps, Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:tas_displays} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, and c) a Dorling cartogram.

@@

```{r tas_displays, width = 10, height = 3, fig.cap = "The three displays show alternative maps of the Statistical Areas at Level 2 in the Australian state of Tasmania. These alternatives show the (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania."}
library(gridExtra)
grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
```

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in Figure \ref{fig:tas_displays}a) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. The algorithm uses the geographic shape of the areas and iterates toward sizing the areas to represent the population. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS].

@@

The non-contiguous cartogram in Figure \ref{fig:tas_displays}b) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Hobart may look reasonable but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:tas_displays}c) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in Figure \ref{fig:tas_displays} a), or b). This display also highlights that the approximation of equal population is not easy to fulfil, as there is some disparity in the sizes of the circles.

@@

The result of the algorithm on rates of thyroid cancer for females in each of the SA2 areas of Australia is shown in Figure \ref{fig:hexmap}.
The difference to the choropleth map in Figure \ref{fig:choro} are clear. Now the thyroid cancer incidence in the densely populated areas in Sydney, Brisbane and Perth are visible. Interestingly there is no clear divide between rural and urban SA2 areas as many rural areas, and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis for females. The hexagon tile map display provides a more accurate representation of the spatial distribution of thyroid cancer incidence across Australia.

When the algorithm is applied to the complete set of all Australian Statistical Areas at Level 2 it results in the hexagon tile map display shown in Figure \ref{fig:hexmap}. This display highlights the density of Australian capital cities, as it draws attention to the many communities in Sydney, Melbourne and Hobart. This display also highlights the disparity in the burden of thyroid cancer for females in the communities of these cities. There are several collections of red hexagons in Sydney that represent the communities with much higher rates of diagnosis than the Australian average. Brisbane also experiences higher than average rates of diagnosis, but has more orange than red. The females in the cities of Adelaide and Perth  show much lower rates of diagnosis.

@@

There are several key steps needed to create the hexagon tile map as described in the flow chart in Figure \ref{fig:sugarbag_flow}. First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart, represented in Figure \ref{fig:filterprocess}c and d as a white cross.

@@

The set of possible hexagon grid points reduces by one after each centroid is allocated, then only the grid points that have not yet been selected are considered. Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon. This is demonstrated in Figure \ref{fig:filterprocess}c and d by the black hexagons that represent the seven closest polygons to the capital city of Hobart. As the allocation process begins for the eighth closest centroid there are seven unavailable hexagon locations.

@@

The `width` parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \ref{fig:filterprocess}c. This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

@@

<!-- Tasmania example has been removed -->
<!-- A complete hexagon tile map of Australia is created by applying the algorithm steps to each centroid. The hexagon tile map is used to visualise the Australian Cancer Atlas data [@TACA]. Two views of the same data are produced by filling according to the thyroid Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in Figure \ref{fig:hexmap} shows the group of blue areas in the Hobart CBD more prominently in the hexagon tile map. The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the rural areas. -->

@@

<!-- Summarise paper, usability -->
A choropleth map display is the current common practice for spatial data visualisation. As the choropleth map display relies on the geographic shapes, it does not effectively present the density of the population in small geographic areas. The common solution to enhance the communication of spatial distributions is to incorporate the population values as a cartogram or another alternative map display. Many alternative map displays do not work well for Australia due to the large rural geographic space between the densely populated capital cities. The administrative boundaries may also distract from the statistics communicated using colour due to the small size of some areas.
This paper provides an algorithm for Australia called the hexagon tile map.

<!-- implementation -->
Alternative maps highlight the value of the statistics across the geographic units. These methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem can be using equally sized areas, with neighbourhood boundary connections. The `sugarbag` package for `R` creates tessellated hexagon tile maps by automating an algorithm to design these displays. The Australian application preserves the spatial relationships, emphasising capital cities.

These hexagons equally represent each area. However, similar to the choropleth map display the tessellation does not allow the size of the hexagons to represent another variable. The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. It works on the assumption that viewers can identify their neighbourhoods using the relative directional relationships but this can be aided by the use of animation. With careful consideration of the choropleth map, the small geographic inner city areas may have been noticed by viewers, but the hexagon tile map display emphasises them. The communities in northern Tasmania and the Northern territory do not draw attention because of their size as in the choropleth, but their colour is still noticeably below average when contrasted with the hexagons in New South Wales.

This algorithm should  be general enough to apply to other countries and problems

<!-- why hexagon tile map addresses deficiencies -->

<!-- The Dorling cartogram was the most influential alternative map display for the design of the hexagon tile map. The use of a hexagon shifts the focus of the visualisation from the shape of SA2s to the distribution effects of a disease on communities. It is not impeded by the irregular and unusual shapes created by a contiguous cartogram and the immense white space created on the non-contiguous cartogram. Hexagons can emphasise the relationship between neighbouring communities by allowing their sides to touch. The tessellation created by adjoining hexagons for densely populated areas is helpful as it shows that most communities are not islands onto themselves, but are intrinsically connected to those around them. As seen in the Dorling cartogram in figure \ref{fig:tas_displays}c), the tessellation that allows boundary connections was not imposed for rural communities, where centroid of the geographic unit was located far from others, a feature that allows distant but dense populations of the coastal towns of southern Tasmania to be recognised. -->

<!-- , limitiations, where it might not work -->

This algorithm has only been tested using single countries, and does not consider definite borders of countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism to protect the borders and ensure centroids are placed within the geographic borders of a country.

<!-- future -->
It would be also possible to take a logarithmic function in the filter steps for selection of a hexagon grid point. A direct angle is currently used, but this may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location, before increasing the width of the angle used to filter the hexagons.
This algorithm is an effective start to creating hexagon tile maps for many geographic units.

>>>>

---
title: A Hexagon Tile Map Algorithm for Displaying Spatial Data
author:
  # see ?rjournal_article for more information
  - name: Stephanie Kobakian
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  stephanie.kobakian@monash.edu 
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
  - name: Earl Duncan
    affiliation: Queensland University of Technology
    address:
    - School of Mathematical Sciences
    email:  earl.duncan@qut.edu.au
abstract: >
  Spatial distributions have been presented on alternative representations of geography for many years. In modern times, interactivity and animation have allowed alternative displays to play a larger role. Alternative representations have been popularised by online news sites, and digital atlases with a focus on public consumption. Applications are increasingly widespread, especially in the areas of disease mapping, and election results. The algorithm presented creates a display that uses tessellated hexagons to represent a set of spatial polygons. It allocates these hexagon in a manner that preserves the spatial relationship of the geographic units, in light of their positions to points of interest. The display showcases spatial distributions, by emphasising the small geographical regions that are often difficult to locate on geographic maps. 
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: kobakian-cook.bib
output: rticles::rjournal_article
editor_options: 
  chunk_output_type: console
---

@@

Over time, the density of living in major cities has increased as residents have gathered to live near urban areas [@ACTUC]. The division of these residents into approximately equal population areas results in dramatically different square meterage of administrative geographic units, such as states or electorates. In a choropleth map display, the geographic units are coloured to represent the value of the statistic for each unit [@EI]. This can cause some values to be emphasised over others, and allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:choro} is a choropleth map that uses colour to display the estimated standardised incidence ratios (SIRs) of thyroid cancer for females, in each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s on the east coast around Brisbane and in northern New South Wales stand out as more orange and red. However, this display neglects the vast amount of Australia residents living in the densely populated capital cities.

@@

The solutions to this visualisation problem begin with the geography. Alternative maps can be created to shift the focus from land area and shape to the value of the statistics [@ACCAC]. Each style of cartogram applies a different transformation to the geographic areas, to highlight the values of the statistic of interest. Alternative maps can result in a distortion of the map space to represent features of the distribution across the areas [@ACCAC] as the statistic of interest is used to determine the cartogram layout.

Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing the population in each administrative area [@TAAM]. This acknowledges that the number of residents can be different and also recognises that each area, or person within it is equally important.

This paper contains a discussion of Existing Mapping Practices. This is followed by details of the Algorithm. The implementation section discusses the algorithm in the `sugarbag` package. Finally, the use and benefits of animation are presented.

@@

<!-- choropleth -->
The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC]. The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS].

Administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. Each SA2 [@abs2011] was designed to represent a community. This collection of communities presents an opportunity for explicit map transformations to improve communication of spatial distributions [@CBATCC]. In figure \ref{fig:tas_displays} the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

There are several established alternative visualisation methods. Tile maps, rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:tas_displays} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, and c) a Dorling cartogram.

@@

```{r tas_displays, width = 10, height = 3, fig.cap = "The three displays show alternative maps of the Australian state of Tasmania. These alternatives show the (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania."}
library(gridExtra)
grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
```

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The connected communities can be recognised by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:tas_displays}a) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. The algorithm uses the geographic shape of the areas and iterates toward sizing the areas to represent the population. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS].

@@

The non-contiguous cartogram in figure \ref{fig:tas_displays}b) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of background space can be meaningful in non-contiguous cartograms [@ECGC]. The size disparity between rural areas and urban areas result in reasonable display within the south eastern city of Hobart, but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:tas_displays}c) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in figure \ref{fig:tas_displays} a), or b). This display also highlights that the approximation of equal population is not easy to fulfil, as there is some disparity in the sizes of the circles.

@@

The result of the algorithm on rates of thyroid cancer for females in each of the SA2 areas of Australia is shown in figure \ref{fig:hexmap}.
The difference to the choropleth map in figure \ref{fig:choro} are clear. Now the thyroid cancer incidence in the densely populated areas in Sydney, Brisbane and Perth are visible. Interestingly there is no clear divide between rural and urban SA2 areas as many rural areas, and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis for females. The hexagon tile map display provides a more accurate representation of the spatial distribution of thyroid cancer incidence across Australia.

When the algorithm is applied to the complete set of all Australian Statistical Areas at Level 2 it results in the hexagon tile map display shown in figure \ref{fig:hexmap}. This display highlights the density of Australian capital cities, as it draws attention to the many communities in Sydney, Melbourne and Hobart. This display also highlights the disparity in the burden of thyroid cancer for females in the communities of these cities. There are several collections of red hexagons in Sydney that represent the communities with much higher rates of diagnosis than the Australian average. Brisbane also experiences higher than average rates of diagnosis, but has more orange than red. The females in the cities of Adelaide and Perth  show much lower rates of diagnosis.

@@

There are several key steps needed to create the hexagon tile map as described in the flow chart in figure \ref{fig:sugarbag_flow}. First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart, represented in figure \ref{fig:filterprocess}c and d as a white cross.

@@

The set of possible hexagon grid points reduces by one after each centroid is allocated, then only the grid points that have not yet been selected are considered. Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon. This is demonstrated in figure \ref{fig:filterprocess}c and d by the black hexagons that represent the seven closest polygons to the capital city of Hobart. As the allocation process begins for the eighth closest centroid there are seven unavailable hexagon locations.

@@

The `width` parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in figure \ref{fig:filterprocess}c. This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

@@

<!-- Tasmania example has been removed -->
<!-- A complete hexagon tile map of Australia is created by applying the algorithm steps to each centroid. The hexagon tile map is used to visualise the Australian Cancer Atlas data [@TACA]. Two views of the same data are produced by filling according to the thyroid Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in figure \ref{fig:hexmap} shows the group of blue areas in the Hobart CBD more prominently in the hexagon tile map. The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the rural areas. -->

@@

This paper provides an algorithm for Australia called the hexagon tile map.
It also describes how this algorithm addresses the potential issues found with contiguous, non-contiguous and Dorling cartograms when applied to Australia. It achieves a display that effectively presents the density of the population in small geographic areas and de emphasises the large rural geographic space between the densely populated capital cities.

<!-- implementation -->
The hexagon tile map display acknowledges that the amount of residents can differ but each administrative area is equally important. This solution uses equally sized areas, and maintain neighbourhood boundary connections. The `sugarbag` package for `R` automates the creation of tessellated hexagon tile maps by providing an algorithm to design these displays. The Australian application preserves spatial relationships, and emphasises capital cities. 
However, similar to the choropleth map display the tessellation does not allow the size of the hexagons to represent another variable. The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. With careful consideration of the choropleth map, the small geographic inner city areas may have been noticed by viewers, but the hexagon tile map display emphasises them. The communities in northern Tasmania and the Northern territory do not draw attention because of their size as in the choropleth, but their colour is still noticeably below average when contrasted with the hexagons in New South Wales.

<!-- limitiations, where it might not work -->

This algorithm should be general enough to apply to other countries and related problems. The algorithm has only been tested using single countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism implemented to protect the borders and ensure centroids are placed within the geographic borders of a country.
The algorithm is also currently limited to representing each geographic area with only one hexagon. Future developments may address these limitations, and refine the filter steps for final selection of a hexagon grid point for each centroid. A direct angle is currently implemented, but may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location.
This algorithm is an effective start to creating hexagon tile maps for many geographic units. This contributes to an extensive body of work that encourages the use of alternative map displays for statistical displays of spatial data.

### 2020-11-18 kobakian-cook.tex
Earl's comments up to Algorithm section

\abstract{%
The algorithm presented creates a display that uses tessellated hexagons
to represent a set of spatial polygons. It allocates these hexagon in a
manner that preserves the spatial relationship of the geographic units,
in light of their positions to points of interest. The display showcases
spatial distributions, by emphasising the small geographical regions
that are often difficult to locate on geographic maps. Spatial
distributions have been presented on alternative representations of
geography for many years. In modern times, interactivity and animation
have allowed alternative displays to play a larger role. Alternative
representations have been popularised by online news sites, and atlas
websites with a focus on public consumption. Applications are
increasingly widespread, especially in the areas of disease mapping, and
election results.
}

@@

Over time the density of living in major cities has increased as
residents have gathered to live near urban areas \citep{ACTUC}. The
division of these resident into approximately equal population areas
results in dramatically different square meterage of administrative
geographic units, such as states or electorates. In a choropleth map
display the geographic units are filled with colour to represent the
value of the statistic for each unit \citep{EI}. This can allow unequal
emphasis on the statistical value of each of the geographic units. This
allows choropleth map displays to misrepresent the spatial distributions
of human related statistics due to area-size bias \citep{BCM}. Figure
\ref{fig:choro} is a choropleth map that uses colour to display the
estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for
females, in each of the Statistical Areas at Level 2 (SA2) used by the
\citet{abs2011}. The Australian choropleth map display draws attention
to the expanse of dark and light blue areas across the rural communities
in all states. The SA2s around Brisbane and in northern New South Wales
stand out as more orange and red. However, this display neglects the
vast amount of Australia residents living in the densely populated
capital cities.

@@

The solutions to this visualisation problem begin with the geography.
Alternative maps shift the focus from land area and shape, to the value
of the statistics in a collection of areas \citep{ACCAC}. Each style of
cartogram applies a different transformation to the geographic areas, to
highlight the values of the statistic of interest. These displays result
in a distortion of the map space to represent features of the
distribution across the areas \citep{ACCAC} as the statistic of interest
is used to determine the cartogram layout.

Alternative mapping methods allow increased understanding of the spatial
distribution of a variable across the population, by fairly representing
the population in each administrative area \citep{TAAM}. This
acknowledges that the number of residents can be different but
recognises that each area, or person within it is equally important.

This paper contains a discussion of Existing Mapping Practices in
Section 2. This is followed by details of the Algorithm in Section 3.
Section 4 describes the implementation of the algorithm in the
\texttt{sugarbag} package. Section 5 outlines the use and benefits of
animation.

@@

A choropleth map display is the current common practice for spatial data
visualisation. As the choropleth map display relies on the geographic
shapes, it does not effectively present the density of the population in
small geographic areas. The common solution to enhance the communication
of spatial distributions is to incorporate the population values as a
cartogram or another alternative map display. Many alternative map
displays do not work well for Australia due to the large rural
geographic space between the densely populated capital cities. The
administrative boundaries may also distract from the statistics
communicated using colour due to the small size of some areas. This
paper provides an algorithm for Australia called the hexagon tile map.

Alternative maps highlight the value of the statistics across the
geographic units. These methods allow increased understanding of the
spatial distribution of a variable across the population, by fairly
representing each administrative area. This acknowledges that the amount
of residents can be different but recognises that each population area
is equally important. The solution to this visualisation problem can be
using equally sized areas, with neighbourhood boundary connections. The
\texttt{sugarbag} package for \texttt{R} creates tessellated hexagon
tile maps by automating an algorithm to design these displays. The
Australian application preserves the spatial relationships, emphasising
capital cities.

These hexagons equally represent each area. However, similar to the
choropleth map display the tessellation does not allow the size of the
hexagons to represent another variable. The algorithm is heavily
dependent on the focal points used, as this determines the order of
allocation. It works on the assumption that viewers can identify their
neighbourhoods using the relative directional relationships but this can
be aided by the use of animation. With careful consideration of the
choropleth map, the small geographic inner city areas may have been
noticed by viewers, but the hexagon tile map display emphasises them.
The communities in northern Tasmania and the Northern territory do not
draw attention because of their size as in the choropleth, but their
colour is still noticeably below average when contrasted with the
hexagons in New South Wales.

This algorithm should be general enough to apply to other countries and
problems

This algorithm has only been tested using single countries, and does not
consider definite borders of countries. While the buffer allows
extension beyond the furthest centroids, there is no mechanism to
protect the borders and ensure centroids are placed within the
geographic borders of a country.

It would be also possible to take a logarithmic function in the filter
steps for selection of a hexagon grid point. A direct angle is currently
used, but this may narrowly miss appropriate hexagons, a logarithmic
function may help to choose a closer hexagon to the original centroid
location, before increasing the width of the angle used to filter the
hexagons. This algorithm is an effective start to creating hexagon tile
maps for many geographic units.

>>>>

\abstract{%
Spatial distributions have been presented on alternative representations
of geography for many years. In modern times, interactivity and
animation have allowed alternative displays to play a larger role.
Alternative representations have been popularised by online news sites,
and digital atlases with a focus on public consumption. Applications are
increasingly widespread, especially in the areas of disease mapping, and
election results. The algorithm presented creates a display that uses
tessellated hexagons to represent a set of spatial polygons. It
allocates these hexagon in a manner that preserves the spatial
relationship of the geographic units, in light of their positions to
points of interest. The display showcases spatial distributions, by
emphasising the small geographical regions that are often difficult to
locate on geographic maps.
}

@@

Over time, the density of living in major cities has increased as
residents have gathered to live near urban areas \citep{ACTUC}. The
division of these residents into approximately equal population areas
results in dramatically different square meterage of administrative
geographic units, such as states or electorates. In a choropleth map
display, the geographic units are coloured to represent the value of the
statistic for each unit \citep{EI}. This can cause some values to be
emphasised over others, and allows choropleth map displays to
misrepresent the spatial distributions of human related statistics due
to area-size bias \citep{BCM}. Figure \ref{fig:choro} is a choropleth
map that uses colour to display the estimated standardised incidence
ratios (SIRs) of thyroid cancer for females, in each of the Statistical
Areas at Level 2 (SA2) used by the \citet{abs2011}. The Australian
choropleth map display draws attention to the expanse of dark and light
blue areas across the rural communities in all states. The SA2s on the
east coast around Brisbane and in northern New South Wales stand out as
more orange and red. However, this display neglects the vast amount of
Australia residents living in the densely populated capital cities.

@@

The solutions to this visualisation problem begin with the geography.
Alternative maps can be created to shift the focus from land area and
shape to the value of the statistics \citep{ACCAC}. Each style of
cartogram applies a different transformation to the geographic areas, to
highlight the values of the statistic of interest. Alternative maps can
result in a distortion of the map space to represent features of the
distribution across the areas \citep{ACCAC} as the statistic of interest
is used to determine the cartogram layout.

Alternative mapping methods allow increased understanding of the spatial
distribution of a variable across the population, by fairly representing
the population in each administrative area \citep{TAAM}. This
acknowledges that the number of residents can be different and also
recognises that each area, or person within it is equally important.

This paper contains a discussion of Existing Mapping Practices. This is
followed by details of the Algorithm. The implementation section
discusses the algorithm in the \texttt{sugarbag} package. Finally, the
use and benefits of animation are presented.

@@

This paper provides an algorithm for Australia called the hexagon tile
map. It also describes how this algorithm addresses the potential issues
found with contiguous, non-contiguous and Dorling cartograms when
applied to Australia. It achieves a display that effectively presents
the density of the population in small geographic areas and de
emphasises the large rural geographic space between the densely
populated capital cities.

The hexagon tile map display acknowledges that the amount of residents
can differ but each administrative area is equally important. This
solution uses equally sized areas, and maintain neighbourhood boundary
connections. The \texttt{sugarbag} package for \texttt{R} automates the
creation of tessellated hexagon tile maps by providing an algorithm to
design these displays. The Australian application preserves spatial
relationships, and emphasises capital cities. However, similar to the
choropleth map display the tessellation does not allow the size of the
hexagons to represent another variable. The algorithm is heavily
dependent on the focal points used, as this determines the order of
allocation. With careful consideration of the choropleth map, the small
geographic inner city areas may have been noticed by viewers, but the
hexagon tile map display emphasises them. The communities in northern
Tasmania and the Northern territory do not draw attention because of
their size as in the choropleth, but their colour is still noticeably
below average when contrasted with the hexagons in New South Wales.

This algorithm should be general enough to apply to other countries and
related problems. The algorithm has only been tested using single
countries. While the buffer allows extension beyond the furthest
centroids, there is no mechanism implemented to protect the borders and
ensure centroids are placed within the geographic borders of a country.
The algorithm is also currently limited to representing each geographic
area with only one hexagon. Future developments may address these
limitations, and refine the filter steps for final selection of a
hexagon grid point for each centroid. A direct angle is currently
implemented, but may narrowly miss appropriate hexagons, a logarithmic
function may help to choose a closer hexagon to the original centroid
location. This algorithm is an effective start to creating hexagon tile
maps for many geographic units. This contributes to an extensive body of
work that encourages the use of alternative map displays for statistical
displays of spatial data.
